---
title: "RWorksheet_Obas#4c"
author: "Barbie Joy Obas"
date: "2023-11-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1.a Show your solutions on how to import a csv file into the environment
```{r}
mpg <- read.csv(file= "mpg.csv")
mpg
```

#1.b Which variables from mpg dataset are categorical?
```{r}
mpg_Var <- names(sapply(mpg, function(x) is.factor(x) || is.character(x)))
mpg_Var
# The variables from mpg dataset are categorical are manufacturer, model, displ, 
# year, cyl, trans, drv, cty, hwy, fl, class
```

#1.c Which are continuous variables?
```{r}
continous <- sapply(mpg, is.numeric)
cont_vars <- names(mpg)[continous]
cont_vars

# The continuous variables are displ, year, cyl, cty, and hwy
```

#2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.
```{r}
library(dplyr)

  mostmodels_man <- mpg %>%
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models)) %>%
  head(1)


  mostmodels_var <- mpg %>%
  group_by(model) %>%
  summarise(num_var = n()) %>%
  arrange(desc(num_var)) %>%
  head(1)


cat("Manufacturer with the most models:", mostmodels_man $manufacturer, "\n")
cat("Model with the most variations:", mostmodels_var$model, "\n")

#Manufacturer with the most models is toyota 
#Model with the most variations is caravan 2wd 
```

#2.a Group the manufacturers and find the unique models. Show your codes and result
```{r}

library(dplyr)

manufacturersModels <- data.frame(Manufacturer = mpg$manufacturer, Model = mpg$model)
manufacturersModels

uniqueMods <- unique(manufacturersModels)
uniqueMods

uniqueModsFactor <- factoredManufacturer <- as.factor(uniqueMods$Manufacturer)

```

#2b Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
library(ggplot2)

library(dplyr)

uniquePlot <- plot(as.factor(factoredManufacturer),
                 main = "Unique Models of Manufacturers",
                 xlab = "Manufacturer", 
                 ylab = "Number of Unique Models",
                 cex.names = 0.8, las = 2)


uniqueCount <- uniqueMods %>% 
  count(uniqueMods$Manufacturer)
uniqueCount

ggplot(uniqueCount, aes(x = `uniqueMods$Manufacturer`, y = n)) +
geom_bar(stat = "identity", fill = "pink") + 
  labs(x = "Manufacturer", y = "Number of Unique Models")

```

#2.1 Same dataset will be used. You are going to show the relationship of the model and the manufacturer.

```{r}
library(ggplot2)

ggplot(mpg, aes(x = manufacturer, fill = model)) +
  geom_bar() +
  labs(title = "Relationship of Model and Manufacturer",
       x = "Manufacturer",
       y = "Count",
       fill = "Model") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#2.1.a What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point()


#The "model" variable is represented by the x-axis. The variable "manufacturer" is represented by the y-axis.Every point corresponds to a unique combination of a car model and its manufacturer.


#To add points to the plot, utilize the geom_point() method. You will see a point for every combination of model and manufacturer because there are two categorical variables (model and manufacturer). The points are dispersed throughout the plot.

```

#2.1.b For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
ggplot(mpg,aes(x = model, y = manufacturer,color = class)) + geom_point()

#This data is helpful for assessing the quantity of models per manufacturer. 

#To heighten its informativeness, I could introduce color-coded points reflecting the class variable and incorporate labels indicating the cty variable for each point.
```

#3.Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results

```{r}
top_20 <- head(mpg,20)

topPlot20 <- ggplot(top_20, aes(x = model, y = year)) + geom_point() + labs(x = "Model", y = "Year") 
topPlot20

```

#4.Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result


```{r}
library(dplyr)

modelCarCount <- mpg %>% 
  group_by(model) %>%
  summarize(numberOfCars = n())

modelCarCount

```

#4.a Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.

```{r}
obs20 <- head(modelCarCount, 20)

top_20 <- ggplot(obs20, aes(x = model, y = numberOfCars, fill = model)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Number of Cars per Model", 
       x = "Model" , 
       y = "Number of cars")

top_20
```
#4.b  Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}

flippedTop20 <- ggplot(obs20, aes(x = model, y = numberOfCars, fill = model)) + 
  geom_bar(stat = "identity") + labs(title = "Number of Cars per Model", x = "Model" ,
                                     y = "Number of Cars")+ coord_flip()

flippedTop20


```

#5.Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

```{r}
cylDisplayPlot <- ggplot(mpg, aes(x = cyl, y = displ, color = displ)) + geom_point() + labs(title = "Relationship between No.of Cylinders and Engine Displacement",x = "Number of Cylinders",y = "Engine Displacement")

cylDisplayPlot

```

#5.a How would you describe its relationship? Show the codes and its result.

```{r}
#This will produce a scatter plot illustrating how the number of cylinders relates to engine displacement. Each point's color indicates its specific engine displacement.

#As cylinder count rises, engine displacement generally follows suit. This indicates a trend where larger engines with more cylinders tend to have increased displacement.
```

